---
title: "Seuqntial Bayes Factor Analysis"
author: "郑元瑞"
date: "2023-01-22"
output: html_document
---

## Import packages and data

```{r import packages and info, message=FALSE}
library(tidyverse) # data wrangling, plotting
library(brms) # Bayesian analysis
library(rstan) # Bayesian analysis
library(tidyverse)
library(here)
library(hypr)
library(bridgesampling)
library(BayesFactor)
library(here)
# devtools::install_github("RobinHankin/Brobdingnag")
# update.packages("dbplyr")
set.seed(123)
options(mc.cores = parallel::detectCores())
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
R.version$version.string
```

### Import four dataset and preprocessing

-   df.agg1 is aggregated data for experiment 1
-   df.agg2 is aggregated data for experiment 2
-   dat1 is the reaction time data without aggregating for experiment 1
-   dat2 is the reaction time data without aggregating for experiment 2

```{r import data, message=FALSE}
df.agg1 <- read_csv(here("1_Data", "2_Postpro_data", "Exp1_RT_Agg.csv"))
dat1 <- read_csv((here("1_Data", "2_Postpro_data", "Exp1_RT.csv")))
df.agg2 <- read_csv(here("1_Data", "2_Postpro_data", "Exp2_RT_Agg.csv"))
dat1 <- read_csv((here("1_Data", "2_Postpro_data", "Exp1_RT.csv")))
dat2 <- read_csv((here("1_Data", "2_Postpro_data", "Exp2_RT.csv")))
```

### Data Preprocessing

rename the df.agg1 and selected the needed columns

for the **seq** variable

-   The I represents image_first in **seq** variable, which means in the experiment, the image is presented before the text.

-   The W represents word_first in **seq** variable, which means in the experiment, the text is presented before the image

-   The S represents simultaneous in **seq** variable, which means in the experiment, the text and the image are present simultaneously

#### Aggregated data
```{r}
df.agg1 <- df.agg1 %>%#rename the columns
  dplyr::rename(
    "seq" = condition,
    "rt" = rt_m
  ) %>%
  dplyr::mutate(seq = dplyr::case_when(
    seq == "image_first" ~ "I",
    seq == "word_first" ~ "W",
    seq == "simultaneous" ~ "S"
  )) %>%
  dplyr::select(subj_idx, valence, matchness, seq, rt) %>% 
  dplyr::mutate(exp = "exp1")
df.agg1$subj_idx <- as.factor(df.agg1$subj_idx)
head(df.agg1)
```

```{r}
df.agg2 <- df.agg2 %>%
  dplyr::rename(
    "seq" = condition,
    "rt" = rt_m
  ) %>%
  dplyr::mutate(seq = dplyr::case_when(
    seq == "image_first" ~ "I",
    seq == "word_first" ~ "W",
    seq == "simultaneous" ~ "S"
  )) %>%
  dplyr::select(subj_idx, valence, matchness, seq, rt) %>% 
  dplyr::mutate(exp = "exp2")
df.agg2$subj_idx <- as.factor(df.agg2$subj_idx)
head(df.agg2)
```

transform the long data to wide for analysing data in JASP
```{r}
df.agg1_wide <- df.agg1 %>%
  tidyr::pivot_wider(
    id_cols = "subj_idx",
    names_from = c(
      valence, matchness,
      seq
    ),
    values_from = rt
  )
df.agg2_wide <- df.agg2 %>%
  tidyr::pivot_wider(
    id_cols = "subj_idx",
    names_from = c(
      valence, matchness,
      seq
    ),
    values_from = rt
  )
outpath_v <- here::here("1_Data", "2_Postpro_data")
# readr::write_csv(df.agg2_wide, paste(outpath_v, "Exp2_RT_agg_wide.csv", sep = "/"))
# readr::write_csv(df.agg1_wide, paste(outpath_v, "Exp1_RT_agg_wide.csv", sep = "/"))
```

#### unaggregated data
```{r}
df.unagg1 <- dat1 %>%
  dplyr::rename("seq" = condition) %>%
  dplyr::mutate(seq = dplyr::case_when(
    seq == "image_first" ~ "I",
    seq == "word_first" ~ "W",
    seq == "simultaneous" ~ "S"
  ))
```

```{r}
df.unagg2 <- dat2 %>%
  dplyr::rename("seq" = condition) %>%
  dplyr::mutate(seq = dplyr::case_when(
    seq == "image_first" ~ "I",
    seq == "word_first" ~ "W",
    seq == "simultaneous" ~ "S"
  ))
```

## Basic information about the dependent variable
### the hist plot for aggregated data

```{r}
df.agg1 %>% 
  ggplot(aes(x = rt)) +
  geom_histogram()
```
```{r}
df.agg1 %>% 
  ggplot(aes(x = log(rt))) +
  geom_histogram()
```

```{r}
df.agg2 %>% 
  ggplot(aes(x = rt)) +
  geom_histogram()
```

```{r}
df.agg2 %>% 
  ggplot(aes(x = log(rt))) +
  geom_histogram()
```

### the hist plot for unaggregated data

```{r}
df.unagg1 %>% 
  ggplot(aes(x = rt)) + 
  geom_histogram()
```

```{r}
df.unagg1 %>% 
  ggplot(aes(x = log(rt))) + 
  geom_histogram()
```



```{r}
df.unagg2 %>% 
  ggplot(aes(x = rt)) + 
  geom_histogram()
```
```{r}
df.unagg2 %>% 
  ggplot(aes(x = log(rt))) + 
  geom_histogram()
```
### Apply log transform to the aggregated data

apply log transform to the aggregated data before entering to the  Bayesfactor package
```{r}
df.agg1 <- df.agg1 %>% 
  dplyr::mutate(log_rt = log(rt))
```


```{r}
df.agg2 <- df.agg2 %>% 
  dplyr::mutate(log_rt = log(rt))
```



## Calculate BF value by RMANOVA and HLM

### Using aggregated data and RMANOVA to calculate BF
```{r}
df.agg1$subj_idx <- as.factor(df.agg1$subj_idx)

subjects1 <- sort(unique(df.agg1$subj_idx))

subjects1

stages <- seq(20, 30, by = 10)

stage_n1 = 0

bfdf1 <- data.frame(effect = NA, N = NA, BF = NA, stage = NA, upper_threshold = NA, lower_threshold = NA)

for (i in stages) {
  data1 <- subset(df.agg1, subj_idx %in% head(subjects1, i))
  
  stage_n1 = stage_n1 + 1
  
  message(length(unique(data1$subj_idx)), " subjects")
  
  message("The data collection stage is ", stage_n1)
  
  data1$subj_idx <- as.factor(data1$subj_idx)
  
  bayesfactors1 <- BayesFactor::generalTestBF(
   log_rt ~ valence * matchness * seq * subj_idx - subj_idx:valence:matchness:seq,
     data = data.frame(data1),
     whichRandom = "subj_idx",
     neverExclude = "subj_idx",
     whichModels = "top"
  )

  summary(bayesfactors1)

  df_output1 <- as.data.frame(bayesfactors1)


  bf_3wayinx1 <- data.frame(effect = "matchness:sequence:valence", N = i,
                           BF = 1/df_output1[1, 1], stage = stage_n1,
                           upper_threshold = 10, lower_threshold = 1/10)


  bf_msinx1 <- data.frame(effect = "matchness:sequence", N = i,
                         BF = 1/df_output1[2, 1], stage = stage_n1,
                         upper_threshold = 6, lower_threshold = 1/6)


  bf_svinx1 <- data.frame(effect = "sequence:valence", N = i,
                         BF = 1/df_output1[3, 1], stage = stage_n1,
                         upper_threshold = 6, lower_threshold = 1/6)


  bf_mvinx1 <- data.frame(effect = "matchness:valence", N = i,
                         BF = 1/df_output1[4, 1], stage = stage_n1,
                         upper_threshold = 6, lower_threshold = 1/6)


  bf_smain1 <- data.frame(effect = "sequence", N = i,
                         BF = 1/df_output1[5, 1], stage = stage_n1,
                         upper_threshold = 10, lower_threshold = 1/10)


  bfdf1 <- rbind(bfdf1, bf_3wayinx1, bf_msinx1, bf_svinx1, bf_mvinx1, bf_smain1)
  
}


```

```{r}

bfdf1 <- na.omit(bfdf1)

readr::write_csv(bfdf1, paste(here::here("2_Analysis", "2_Result"), "BF_exp1.csv", sep = "/"))

bfdf1$stage <- as.factor(bfdf1$stage)

bfdf1 %>% 
  ggplot(aes(x = N, y = log(BF))) + 
  geom_point(aes(color = factor(stage))) + 
  geom_line() + 
  scale_x_continuous(limits = c(20, 30), breaks = c(20, 30))+
  facet_wrap(~effect, scales = "free_y") + 
  geom_hline(aes(yintercept = log(upper_threshold)), linetype=2) + 
  geom_hline(aes(yintercept = log(lower_threshold)), linetype=2) 
```


```{r}
df.agg2$subj_idx <- as.factor(df.agg2$subj_idx)

subjects2 <- sort(unique(df.agg2$subj_idx))

subjects2

stages <- seq(20, 30, by = 10)

stage_n2 = 0

bfdf2 <- data.frame(effect = NA, N = NA, BF = NA, stage = NA, upper_threshold = NA, lower_threshold = NA)

for (i in stages) {
  data2 <- subset(df.agg2, subj_idx %in% head(subjects2, i))
  
  data2$subj_idx <- as.factor(data2$subj_idx)
    
  stage_n2 = stage_n2 + 1
  
  message(length(unique(data2$subj_idx)), " subjects")
  
  message("The data collection stage is ", stage_n2)
  
  bayesfactors2 <- BayesFactor::generalTestBF(
   log_rt ~ valence * matchness * seq * subj_idx - subj_idx:valence:matchness:seq,
     data = data.frame(data2),
     whichRandom = "subj_idx",
     neverExclude = "subj_idx",
     whichModels = "top"
  )
  
  summary(bayesfactors2)
  
  df_output2 <- as.data.frame(bayesfactors2)


  bf_3wayinx2 <- data.frame(effect = "matchness:sequence:valence", N = i,
                           BF = 1/df_output2[1, 1], stage = stage_n2, 
                           upper_threshold = 10, lower_threshold = 1/10)

  
  bf_msinx2 <- data.frame(effect = "matchness:sequence", N = i, 
                         BF = 1/df_output2[2, 1], stage = stage_n2, 
                         upper_threshold = 6, lower_threshold = 1/6)

  
  bf_svinx2 <- data.frame(effect = "sequence:valence", N = i, 
                         BF = 1/df_output2[3, 1], stage = stage_n2, 
                         upper_threshold = 6, lower_threshold = 1/6)

  
  bf_mvinx2 <- data.frame(effect = "matchness:valence", N = i, 
                         BF = 1/df_output2[4, 1], stage = stage_n2, 
                         upper_threshold = 6, lower_threshold = 1/6)

  
  bf_smain2 <- data.frame(effect = "sequence", N = i, 
                         BF = 1/df_output2[5, 1], stage = stage_n2, 
                         upper_threshold = 10, lower_threshold = 1/10)

  
  bfdf2 <- rbind(bfdf2, bf_3wayinx2, bf_msinx2, bf_svinx2, bf_mvinx2, bf_smain2)
  
}


```

```{r}

bfdf2 <- na.omit(bfdf2)

readr::write_csv(bfdf2, paste(here::here("2_Analysis", "2_Result"), "BF_exp2.csv", sep = "/"))

bfdf2$stage <- as.factor(bfdf2$stage)

bfdf2 %>% 
  ggplot(aes(x = N, y = log(BF))) + 
  geom_point(aes(color = factor(stage))) + 
  geom_line() + 
  scale_x_continuous(limits = c(20, 30), breaks = c(20, 30))+
  facet_wrap(~effect, scales = "free_y") + 
  geom_hline(aes(yintercept = log(upper_threshold)), linetype=2) + 
  geom_hline(aes(yintercept = log(lower_threshold)), linetype=2) 
```


```{r}
m_inx1 <- brm(rt ~ valence * matchness * seq + (valence * matchness * seq | subj_idx),
  data = df.unagg1,
  family = lognormal(),
  warmup = 2000, iter = 25000, chains = 4,
  control = list(
    adapt_delta = .99,
    max_treedepth = 12
  ),
  save_pars = save_pars(all = TRUE),
  file = "maximal_model_exp1_log25000iter"
)
```


```{r}
m_exin1 <- brm(rt ~ valence + matchness + seq + valence:matchness + valence:seq + matchness:seq + (valence * matchness * seq | subj_idx),
  data = df.unagg1,
  family = lognormal(),
  warmup = 2000, iter = 25000, chains = 4,
  control = list(adapt_delta = .99, max_treedepth = 12),
  save_pars = save_pars(all = TRUE),
  file = "exinx_model_exp1_log25000iter"
)
```
```{r}
rbind(df.agg1, )
```





```{r}
m_inx2 <- brm(rt ~ valence * matchness * seq + (valence * matchness * seq | subj_idx),
  data = df.unagg2,
  family = lognormal(),
  warmup = 2000, iter = 25000, chains = 4,
  control = list(
    adapt_delta = .99,
    max_treedepth = 12
  ),
  save_pars = save_pars(all = TRUE),
  file = "maximal_model_exp2_log25000iter"
)
```

```{r}
fixef(m_inx2)
```



```{r}
m_exin2 <- brm(rt ~ valence + matchness + seq + valence:matchness + valence:seq + matchness:seq + (valence * matchness * seq | subj_idx),
  data = df.unagg2,
  family = lognormal(),
  warmup = 2000, iter = 25000, chains = 4,
  control = list(adapt_delta = .99, max_treedepth = 12),
  save_pars = save_pars(all = TRUE),
  file = "exinx_model_exp2_log25000iter"
)
```

```{r}
fixef(m_exin2)
```

```{r}
m_inx1 <-  readRDS(here("2_Analysis", "2_Result",  "maximal_model_exp1_log25000iter.rds"))
m_exin1 <- readRDS(here("2_Analysis", "2_Result",  "exinx_model_exp1_log25000iter.rds"))
```


```{r}
m_inx2 <-  readRDS(here("2_Analysis", "2_Result",  "maximal_model_exp2_log25000iter.rds"))
m_exin2 <- readRDS(here("2_Analysis", "2_Result",  "exinx_model_exp2_log25000iter.rds"))
```

```{r}
BF_inx1_1 <- bridgesampling::bridge_sampler(m_inx1, silent = T)
BF_exin1_0 <- bridgesampling::bridge_sampler(m_exin1, silent = T)
```

```{r}
BF_inx2_1 <- bridgesampling::bridge_sampler(m_inx2, silent = T)
BF_exin2_0 <- bridgesampling::bridge_sampler(m_exin2, silent = T)
```


```{r}
bf_inx1 <- bridgesampling::bayes_factor(BF_inx1_1, BF_exin1_0)
bf_inx2 <- bridgesampling::bayes_factor(BF_inx2_1, BF_exin2_0)
```

```{r}
bf_inx1
```
```{r}
bf_inx2
```











